---
title: "jlm_wrangling"
author: "Jenn McNeill"
date: "2024-04-08"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r}

library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(outliers)
library(tidyverse)
library(smooth)
library(dplyr)
library(cowplot)

```

Provide information on how the dataset for this analysis were collected (source), the data contained in the dataset (format). Describe how you wrangled/processed your dataset to get the time series object.

Add a table that summarizes your data structure (variables, units, ranges and/or central tendencies, data source if multiple are used, etc.). This table should inserted as a `kable` function in an R chunk. Just show the first 10 rows of your data. Do not include the code used to generate your table.

```{r import data}

getwd()

auser_raw <- read.csv(file="./Data/Raw/Aquifer_Auser.csv", header=TRUE)
#01/01/2006 is where NAs stop in the first column

auser_raw$Date <- dmy(auser_raw$Date) 

auser_longer <- pivot_longer(auser_raw, Depth_to_Groundwater_LT2:Depth_to_Groundwater_DIEC, names_to = "GW_Well", values_to = "Depth")

auser_depths <- auser_longer %>%
  select(Date, GW_Well, Depth)

#plot all depths together
ggplot(auser_depths, aes(x=Date, color=GW_Well))+
  geom_line(aes(y=Depth))+
  scale_x_date(date_breaks = "5 years", date_labels = "%Y", 
               limits = c(as.Date("2002-01-01"), as.Date("2022-12-31")),
               expand = c(0, 0))+
  theme_light()

ggplot(auser_raw)+
  geom_line(aes(x=Date, y=Depth_to_Groundwater_LT2))+
  scale_x_date(date_breaks = "5 years", date_labels = "%Y", 
               limits = c(as.Date("2002-01-01"), as.Date("2022-12-31")),
               expand = c(0, 0))

tail(auser_raw$Depth_to_Groundwater_LT2,30)

ggplot(auser_raw)+
  geom_line(aes(x=Date, y=Depth_to_Groundwater_SAL))+
  scale_x_date(date_breaks = "5 years", date_labels = "%Y", 
               limits = c(as.Date("2007-01-01"), as.Date("2022-12-31")),
               expand = c(0, 0))

tail(auser_raw$Depth_to_Groundwater_SAL,50)

ggplot(auser_raw)+
  geom_line(aes(x=Date, y=Depth_to_Groundwater_PAG))+
  scale_x_date(date_breaks = "5 years", date_labels = "%Y", 
               limits = c(as.Date("2007-01-01"), as.Date("2022-12-31")),
               expand = c(0, 0))

ggplot(auser_raw)+
  geom_line(aes(x=Date, y=Depth_to_Groundwater_CoS))+
  scale_x_date(date_breaks = "5 years", date_labels = "%Y", 
               limits = c(as.Date("2007-01-01"), as.Date("2022-12-31")),
               expand = c(0, 0))

tail(auser_raw$Depth_to_Groundwater_CoS,50)

ggplot(auser_raw)+
  geom_line(aes(x=Date, y=Depth_to_Groundwater_DIEC))+
  scale_x_date(date_breaks = "5 years", date_labels = "%Y", 
               limits = c(as.Date("2007-01-01"), as.Date("2022-12-31")),
               expand = c(0, 0))

ggplot(auser_raw)+
  geom_line(aes(x=Date, y=Depth_to_Groundwater_LT2))+
  scale_x_date(date_breaks = "5 years", date_labels = "%Y", 
               limits = c(as.Date("2002-01-01"), as.Date("2022-12-31")),
               expand = c(0, 0))

#plot all depth to groundwater lines together
#realize that some lines have values of zero
#convert all zeros to NAs
#convert all series to time series objects
#run the ts clean function to interpolate missing values

auser_nas <- auser_raw %>%
  select(Date, Depth_to_Groundwater_LT2:Depth_to_Groundwater_DIEC) %>%
  mutate_at(vars(Depth_to_Groundwater_LT2:Depth_to_Groundwater_DIEC), 
            ~ ifelse(. == 0, NA, .))

ts_LT2 <- ts(auser_nas[,2],start=c(1998,03,05), frequency=365)

```



